---
title: "Verifying QA task decomposition"
author: "PT"
date: "2024-12-31"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
```

# Explaining answers

The live experiment can be found [here](https://polina-tsvilodub.github.io/goal-inferences/experiments/qa_explanations/).
Below we look at descriptive stats over different types of explanations participants generate, when exposed to competitor responses from case studies 2 and 3. The situations and responses are presented in third person. These results are from pilots in contexts where the participants explain the answers to an alien Bo.

```{r, message=FALSE, warning=FALSE}
cs2 <- read_csv("../data/qa_explanations/results_8_cs2_pilot2.csv")
cs3 <- read_csv("../data/qa_explanations/results_9_cs3_pilot1.csv")
```

Below we plot the counts of the different explanations. The explanations were manually classified into the following categories:

* inference: any explanation that made some kind of reference to the questioner. For cs2, this often included explanations saying that the response offers something that is similar to what the questioner asked for. For cs3, it includes both answers generally mentioning that the questioner wants the item, and those explaining functional properties of the items.
* literal: explanations that focused on the "no" part, saying that the item asked for was not available.
* other: unclassifiable explanations.
* similarity: explanations only mentioning the similarity between the target and offered option, but not mentioning the questioner.

```{r}
cs2_main <- cs2 %>% filter(correct_response == "main") %>%
  mutate(experiment = "case study 2")
cs3_main <- cs3 %>% filter(correct_response == "main") %>%
 mutate(experiment = "case study 3")

explanations_pilot <- rbind(cs2_main, cs3_main)

explanations_pilot %>%
  group_by(experiment, category) %>%
  summarise(
    category_count = n()
  ) %>%
  ggplot(., aes(x = category, y = category_count, fill = category)) +
  geom_col() +
  facet_wrap(experiment~.)
```

## Pilot 3

In this pilot for cs 2 materials, we update the instructions on each main trial to include the following: "..., which thoughts would justify mentioning the 'competitor' instead of mentioning something else".

Furthermore, we introduce a more fine-grained categorization of participants' responses:

* *reference to goals* (clearly stating what the questioner wants to do); example:
  * "Netflix is a service that provides movies (along with other content). It is possible that the friend may want to watch a movie so the statement that they have movies may satisfy their need."
  * "This response is justified because gingerbread is similar to cookies in texture and purpose (a sweet snack), and the flight attendant aims to offer something close to what the passenger might enjoy."
  * "Bo, your roommate suggests sunscreen in response to the friend's request for an umbrella since it's a good strategy to avoid sunburn, which is probably the friend's top concern."
* *reference to needs* (includes "being helpful to questioner"); examples:
  * "The assistant clearly answers the question by saying that they don't have what he is looking for but offers an alternative that is more helpful than simply stopping the conversation there (which would have been less helpful). The alternative may or may not be acceptable but is offered just in case"
  * "He responded with a circus show because he sees the woman is looking for a theatre performance for children. They do not have a theatre performance, but the circus show is for children if she is looking for something different."
  * "...another option they might like"
  * "The class that was suggested is close enough to what the student wanted, the one that was not offered."
  * "A ziplock bag could be used instead of a box."
  * "The response is responsible in the sense that she can use the hair dryer to dry her hair."
* *reference to desires / interests / preferences* (of the questioner), more likely mentioning some features 
  * "Firstly since they don't have the item the assistant must say they do not have it. The assistant however wants to be helpful and knows they are interested in Jewellery and therefore offers a plausible alternative that they may or may not be interested in"
* *similarity* (or relatedness / being reasonable) of the items without mentioning the questioner; examples:
  * "Once again, he is simply explaining that what she is asking they do not have, but he is offering an alternative."
  * "Thrillers are closely related to mysteries as they are both exciting and keep the reader guessing about the ending."
* *feature* (mentioning relevant functional dimension); example:
  * "Neither a cheese pizza nor a pepperoni pizza have vegetables on them."
  * "This is a reasonable response because he doesn't not own a broom, however a vacuum is capable of doing the same function as a broom."
* *literal*; example:
  * "The cashier responds by saying what they currently sell which is mystery novels."
* *other* unclassifiable responses

```{r}
cs2_pilot3 <- read_csv("../data/qa_explanations/results_12_cs2_pilot3.csv")

cs2_pilot3_main <- cs2_pilot3 %>% filter(correct_response == "main") %>%
  mutate(experiment = "case study 2 pilot 3")

explanations_pilot_fineCat <- rbind(cs2_main %>% select(-category),
                                    cs3_main %>% select(-category), 
                                    cs2_pilot3_main)

explanations_pilot_fineCat %>%
  filter(fine_category != "other") %>%
  group_by(experiment, fine_category) %>%
  summarise(
    category_count = n()
  ) %>%
  ggplot(., aes(x = fine_category, y = category_count, fill = fine_category)) +
  geom_col() +
  facet_wrap(experiment~.) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))

```

# Inferring goals from questions

The live experiment can be found [here](https://polina-tsvilodub.github.io/goal-inferences/experiments/goal_sampling/).
Below we preprocess and analyse the results for the pilot where participants suggested three possible goals a questioner had in mind, given the context and the question. The context included the three available options (competitor, same category, other category). Again, the participants' task was to explain questioning behavior to an alien Bo. The results are for vignettes from case study 2 only.

```{r, message=FALSE, warning=FALSE}
goals_cs2 <- read_csv("../data/goal_sampling/results_7_cs2_pilot1.csv") %>%
  filter(correct_response == "main")
goals_cs2_long <- goals_cs2 %>%
  pivot_longer(cols = c(answer1, answer2, answer3), names_to = "answer_num", values_to = "sampled_goals")
#goals_cs2_long %>% write_csv("../data/goal_sampling/results_7_cs2_pilot1_long.csv")

goals_cs2_long <- read_csv("../data/goal_sampling/results_7_cs2_pilot1_long.csv")
```

Each of the suggested goals was manually categorized into the following categories:

* target: includes goals stating that the questioner wants the actual target item, or stating the relevant feature of the target (e.g., "wants an iced non-alcoholic beverage").
* superordinate: includes more general goals, which are mostly epistemic (e.g., something like learning what is on the menu).
* subordinate: includes all other more specific goals. Sometimes the distinction between super- and subordinate wasn't 100% clear; some subordinate goals are, e.g., "wants to know what is on the menu to tell a friend about it".

Some spill-over between the options and the produced goals was observed; i.e., some responses assumed that the questioner knows the options, which should not be the case. These were mostly put towards subordinate goals. The next pilot removes the options from context.

```{r}
goals_cs2_long %>%
  group_by(category) %>%
  summarize(category_prop = n() / nrow(goals_cs2_long)) %>%
  ggplot(., aes(x = category, y = category_prop, fill = category)) +
  geom_col()

```

## Pilot 2

In this pilot, we remove the available options from the context. The participants only see the contextual setting and the question.

```{r, message=FALSE, warning=FALSE}
goals_cs2_p2 <- read_csv("../data/goal_sampling/results_10_cs2_pilot2.csv") %>%
  filter(correct_response == "main")
goals_cs2_p2_long <- goals_cs2_p2 %>%
  pivot_longer(cols = c(answer1, answer2, answer3), names_to = "answer_num", values_to = "sampled_goals")
#goals_cs2_p2_long %>% write_csv("../data/goal_sampling/results_10_cs2_pilot2_long.csv")

goals_cs3 <- read_csv("../data/goal_sampling/results_11_cs3_pilot1.csv") %>%
  filter(correct_response == "main")
goals_cs3_long <- goals_cs3 %>%
  pivot_longer(cols = c(answer1, answer2, answer3), names_to = "answer_num", values_to = "sampled_goals")
#goals_cs3_long %>% write_csv("../data/goal_sampling/results_11_cs3_pilot1_long.csv")

goals_cs2_p2_long <- read_csv("../data/goal_sampling/results_10_cs2_pilot2_long.csv") %>%
  mutate(experiment = "case study 2")
goals_cs3_long <- read_csv("../data/goal_sampling/results_11_cs3_pilot1_long.csv") %>%
   mutate(experiment = "case study 3") %>% filter(category != "other")
goals_long <- rbind(goals_cs2_p2_long, goals_cs3_long)
```

The goals were manually annotated into the same three categories as in pilot 1, but included a more fine-grained distinction between "general" goals and "feature"-mentioning goals. The feature-mentioning goals appealed to the relevant feature of the target (e.g., "wants to do martial arts" rather than just saying "wants to do kickboxing"). Generally, compared to the pilot above, the variation in the contents was larger, the subordinate responses more often mentioned social goals (e.g., the question as an excuse to start a conversation with the cashier). 
In case study 3, there were a few particular aspects to be noted: the larger proportion of feature-target responses is partially driven by repetitions of features from the context (e.g., "wants to reach the window"). The distinction between feature- and non-feature subordinate / superordinate goals was a bit more difficult to draw (e.g., it was unclear whether goals like "to avoid cutting or slippery when holding the mirror" were referencing a feature in the same sense as the target feature of a blanket being soft).

```{r}
goals_long_processed <- goals_long %>%
  mutate(category_split = str_split(category, "-")) %>% 
  unnest_wider(category_split, names_sep = "wide") %>%
  rename(level = category_splitwide1, is_feature = category_splitwide2) %>%
  mutate(is_feature = ifelse(is.na(is_feature), "general", "feature"))

goals_long_processed %>%
  group_by(experiment) %>%
  mutate(expt_count = n()) %>%
  group_by(experiment, level, is_feature) %>%
  summarize(category_prop = n() / expt_count) %>%
  unique() %>%
  ggplot(., aes(x = level, y = category_prop, fill = is_feature)) +
  geom_col() +
  facet_wrap(experiment~.)

```