---
title: "Verifying QA task decomposition"
author: "PT"
date: "2024-12-31"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
```

# Explaining answers

The live experiment can be found [here](https://polina-tsvilodub.github.io/goal-inferences/experiments/qa_explanations/).
Below we look at descriptive stats over different types of explanations participants generate, when exposed to competitor responses from case studies 2 and 3. The situations and responses are presented in third person. These results are from pilots in contexts where the participants explain the answers to an alien Bo.

```{r, message=FALSE, warning=FALSE}
cs2 <- read_csv("../data/qa_explanations/results_8_cs2_pilot2.csv")
cs3 <- read_csv("../data/qa_explanations/results_9_cs3_pilot1.csv")
```

Below we plot the counts of the different explanations. The explanations were manually classified into the following categories:

* inference: any explanation that made some kind of reference to the questioner. For cs2, this often included explanations saying that the response offers something that is similar to what the questioner asked for. For cs3, it includes both answers generally mentioning that the questioner wants the item, and those explaining functional properties of the items.
* literal: explanations that focused on the "no" part, saying that the item asked for was not available.
* other: unclassifiable explanations.
* similarity: explanations only mentioning the similarity between the target and offered option, but not mentioning the questioner.

```{r}
cs2_main <- cs2 %>% filter(correct_response == "main") %>%
  mutate(experiment = "case study 2")
cs3_main <- cs3 %>% filter(correct_response == "main") %>%
 mutate(experiment = "case study 3")

explanations_pilot <- rbind(cs2_main, cs3_main)

explanations_pilot %>%
  group_by(experiment, category) %>%
  summarise(
    category_count = n()
  ) %>%
  ggplot(., aes(x = category, y = category_count, fill = category)) +
  geom_col() +
  facet_wrap(experiment~.)
```

# Inferring goals from questions

The live experiment can be found [here](https://polina-tsvilodub.github.io/goal-inferences/experiments/goal_sampling/).
Below we preprocess and analyse the results for the pilot where participants suggested three possible goals a questioner had in mind, given the context and the question. The context included the three available options (competitor, same category, other category). Again, the participants' task was to explain questioning behavior to an alien Bo. The results are for vignettes from case study 2 only.

```{r, message=FALSE, warning=FALSE}
goals_cs2 <- read_csv("../data/goal_sampling/results_7_cs2_pilot1.csv") %>%
  filter(correct_response == "main")
goals_cs2_long <- goals_cs2 %>%
  pivot_longer(cols = c(answer1, answer2, answer3), names_to = "answer_num", values_to = "sampled_goals")
#goals_cs2_long %>% write_csv("../data/goal_sampling/results_7_cs2_pilot1_long.csv")

goals_cs2_long <- read_csv("../data/goal_sampling/results_7_cs2_pilot1_long.csv")
```

Each of the suggested goals was manually categorized into the following categories:

* target: includes goals stating that the questioner wants the actual target item, or stating the relevant feature of the target (e.g., "wants an iced non-alcoholic beverage").
* superordinate: includes more general goals, which are mostly epistemic (e.g., something like learning what is on the menu).
* subordinate: includes all other more specific goals. Sometimes the distinction between super- and subordinate wasn't 100% clear; some subordinate goals are, e.g., "wants to know what is on the menu to tell a friend about it".

Some spill-over between the options and the produced goals was observed; i.e., some responses assumed that the questioner knows the options, which should not be the case. These were mostly put towards subordinate goals. The next pilot removes the options from context.

```{r}
goals_cs2_long %>%
  group_by(category) %>%
  summarize(category_prop = n() / nrow(goals_cs2_long)) %>%
  ggplot(., aes(x = category, y = category_prop, fill = category)) +
  geom_col()

```

## Pilot 2

In this pilot, we remove the available options from the context. The participants only see the contextual setting and the question.

```{r, message=FALSE, warning=FALSE}
goals_cs2_p2 <- read_csv("../data/goal_sampling/results_10_cs2_pilot2.csv") %>%
  filter(correct_response == "main")
goals_cs2_p2_long <- goals_cs2_p2 %>%
  pivot_longer(cols = c(answer1, answer2, answer3), names_to = "answer_num", values_to = "sampled_goals")
#goals_cs2_p2_long %>% write_csv("../data/goal_sampling/results_10_cs2_pilot2_long.csv")

goals_cs3 <- read_csv("../data/goal_sampling/results_11_cs3_pilot1.csv") %>%
  filter(correct_response == "main")
goals_cs3_long <- goals_cs3 %>%
  pivot_longer(cols = c(answer1, answer2, answer3), names_to = "answer_num", values_to = "sampled_goals")
#goals_cs3_long %>% write_csv("../data/goal_sampling/results_11_cs3_pilot1_long.csv")

goals_cs2_p2_long <- read_csv("../data/goal_sampling/results_10_cs2_pilot2_long.csv") %>%
  mutate(experiment = "case study 2")
goals_cs3_long <- read_csv("../data/goal_sampling/results_11_cs3_pilot1_long.csv") %>%
   mutate(experiment = "case study 3") %>% filter(category != "other")
goals_long <- rbind(goals_cs2_p2_long, goals_cs3_long)
```

The goals were manually annotated into the same three categories as in pilot 1, but included a more fine-grained distinction between "general" goals and "feature"-mentioning goals. The feature-mentioning goals appealed to the relevant feature of the target (e.g., "wants to do martial arts" rather than just saying "wants to do kickboxing"). Generally, compared to the pilot above, the variation in the contents was larger, the subordinate responses more often mentioned social goals (e.g., the question as an excuse to start a conversation with the cashier). 
In case study 3, there were a few particular aspects to be noted: the larger proportion of feature-target responses is partially driven by repetitions of features from the context (e.g., "wants to reach the window"). The distinction between feature- and non-feature subordinate / superordinate goals was a bit more difficult to draw (e.g., it was unclear whether goals like "to avoid cutting or slippery when holding the mirror" were referencing a feature in the same sense as the target feature of a blanket being soft).

```{r}
goals_long_processed <- goals_long %>%
  mutate(category_split = str_split(category, "-")) %>% 
  unnest_wider(category_split, names_sep = "wide") %>%
  rename(level = category_splitwide1, is_feature = category_splitwide2) %>%
  mutate(is_feature = ifelse(is.na(is_feature), "general", "feature"))

goals_long_processed %>%
  group_by(experiment) %>%
  mutate(expt_count = n()) %>%
  group_by(experiment, level, is_feature) %>%
  summarize(category_prop = n() / expt_count) %>%
  unique() %>%
  ggplot(., aes(x = level, y = category_prop, fill = is_feature)) +
  geom_col() +
  facet_wrap(experiment~.)

```